---
title: "Chlorophyll-a colour scale: the SeaWiFS palette"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    fig_width: 7
    fig_height: 7
vignette: >
  %\VignetteIndexEntry{NASA ocean colour}
  \usepackage[utf8]{inputenc}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

We need to plot satellite remote sensing chlorophyll-a data with the NASA palette and a correct scale. 

The R package *palr* contains a function to return the required colours and values, but it contains a large maximum value to account for large possible summary values. 

We specify a maximum range on the values, and plot the scale in two different ways. 

```{r}
maxrange <- c(0.01, 10)
library(palr)
pal <- chl_pal(palette = TRUE)

sb <- which(pal$breaks >= maxrange[1] & pal$breaks <= maxrange[2])

plot(pal$breaks, col = pal$cols, log = "y", ylim = maxrange)
## select values from the NASA legend
abline(h = c(0.01, 0.03, 0.1, 0.3, 1, 3, 10))
```

```{r}
op <- par(fig = c(0.3, 0.7, 0, 1)); 
plot(NA_real_, xlim = c(0, 1), ylim = maxrange, xlab = "", ylab = "", log = "y")
rasterImage(matrix(rev(pal$cols[sb])), 0, 0.01, 1, 10)
abline(h = c(0.01, 0.03, 0.1, 0.3, 1, 3, 10))
par(op)
```

Compare these plots with the official NASA legend here. 

https://oceancolor.gsfc.nasa.gov/l3/



