---
title: "Colour palettes for data in palr"
date: "`r Sys.Date()`"
output: html_vignette
vignette: >
  %\VignetteIndexEntry{Colour palettes}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r, echo = FALSE, message = FALSE}
##knitr::opts_chunk$set(collapse = TRUE, comment = "")
library(palr)
library(raster)
```

# palr

Colours can be frustrating to get just right in R. The palr package provides simple palette functions with standard
colour schemes matched to real data values. 

There are three main ways of working with palr palette functions.  

* `pal(n)` return n colours from the palette
* `pal(data)` return the right colour for values in `data`
* `pal(palette = TRUE)` return the entire palette, with colours `cols` and intervals `breaks`

## Examples

Here we show examples of the use of palr functions. 

The `oisst` data set is a subset the NOAA 1/4° daily Optimum Interpolation Sea Surface Temperature (Reynolds, 2007) for the `r format(getZ(oisst), "%d %B %Y")` obtained from the National Oceanic and Atmospheric Administration (NOAA). 

```{r}
library(raster)
library(palr)
data(oisst)
plot(oisst)
```

The default plot colours uses a setting provided by the raster package, but we have SST data in degrees Celsius so we can use the `sstPal` function to give specific colours for particular temperatures. The full range of the temperatures is shown on the plot legend, even though our data only has values in the range `r format(cellStats(oisst, range), digits = 2)`. 

```{r}
sstcols <- sstPal(palette = TRUE)
plot(oisst, col = sstcols$col, zlim = range(sstcols$breaks))
```

Because we have the palette colours and data in an absolute palette we can also plot other data correctly to scale. 

Create a aggregated version of the SST data set and turn it into polygons. 

```{r}
smsst <- focal(aggregate(oisst, fact = 5, fun = mean), matrix(1, 3), fun = median, na.rm = FALSE)
psst <- rasterToPolygons(smsst, na.rm = FALSE)
layout(matrix(1:2, ncol = 2))
plot(oisst, col = sstcols$col, zlim = range(sstcols$breaks))
plot(oisst, col = "transparent", legend = FALSE)## plot twice to get exact alignment
plot(psst, col = sstPal(psst$layer), add = TRUE)
temps <- rev(c(0, 2, 4, 6, 8, 10, 12, 14, 16, 18, 20))
op <- par(xpd = NA); legend("right", legend = temps, fill = sstPal(temps), inset = -0.2, cex = 0.65)
par(op)
```

## References

Reynolds, R. W., T. M. Smith, C. Liu, D. B. Chelton, K. S. Casey, and M. G. Schlax, 2007: Daily high-resolution-blended analyses for sea surface temperature. Journal of Climate, 20, 5473–5496.

## Appendix: notes on data sets

Full ncdump metadata from the data sets can be obtained with the following. 

```{r}
metadata(oisst)$filename
writeLines(metadata(oisst)$ncdump)
```

